% Lab 3 — compact version (full docs: docs/web/docs/labs/lab3/)
\section{Lab 3: Incident Response and Log Forensics --- HELIX Systems}

Lab~3 inverts the student's role: instead of attacking, they act as a security
analyst responding to an active incident at \textit{HELIX Systems}. The lab
environment comes pre-staged with evidence of a simulated intrusion --- brute-force
SSH attempts, a successful login, privilege escalation via \texttt{sudo}
misconfiguration, backdoor account creation, and lateral movement to a database
server. Students must reconstruct the attack timeline from system logs
(\texttt{auth.log}, \texttt{bash\_history}) and identify the attacker's actions
and impact, following a methodology aligned with NIST SP~800-61~\cite{nist_sp80061}.
Full evidence walkthrough, account table, and attack timeline are in the \href{https://TheEgea.github.io/TFG/annexos/app-d/}{project online documentation}; interactive documentation at \path{docs/web/docs/labs/lab3/}.

\subsubsection*{Network Addressing}

\begin{table}[H]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{@{}lllll@{}}
\toprule
\textbf{Node} & \textbf{Role} & \textbf{Interface} & \textbf{IP} & \textbf{Network} \\
\midrule
Parrot-OS    & Analyst station   & eth0  & 10.0.4.10   & 10.0.4.0/24 (WAN-sim) \\
pfSense-LAB3 & Firewall/DNAT     & WAN   & 10.0.4.1    & 10.0.4.0/24            \\
pfSense-LAB3 & Firewall          & LAN   & 192.168.3.1 & 192.168.3.0/24         \\
VyOS-LAB3    & Internal router   & eth0  & 192.168.3.2 & 192.168.3.0/24         \\
VyOS-LAB3    & Internal router   & eth1  & 10.0.5.1    & 10.0.5.0/24 (corp)     \\
Server-Web   & Compromised web   & eth0  & 10.0.5.10   & 10.0.5.0/24            \\
Server-DB    & Internal DB       & eth0  & 10.0.5.20   & 10.0.5.0/24            \\
\bottomrule
\end{tabular}}
\caption{Lab~3 IP addressing scheme}
\label{tab:lab3-ip}
\end{table}

\subsubsection*{MITRE ATT\&CK Coverage}

\begin{table}[H]
\centering
\small
\begin{tabularx}{\textwidth}{llX}
\toprule
\textbf{MITRE ID} & \textbf{Technique} & \textbf{Evidence in lab} \\
\midrule
T1110.001 & Brute Force: Password Guessing & \texttt{auth.log} failed attempts \\
T1078     & Valid Accounts                 & \texttt{devops} account compromise \\
T1548.003 & Sudo and Sudo Caching          & \texttt{find} NOPASSWD in sudoers \\
T1136.001 & Create Account: Local          & Backdoor \texttt{sysbackup} account \\
T1021.004 & Remote Services: SSH           & Lateral movement to Server-DB \\
T1005     & Data from Local System         & \texttt{/opt/helix/data/} exfil marker \\
\bottomrule
\end{tabularx}
\caption{Lab~3 MITRE ATT\&CK technique coverage}
\label{tab:lab3-mitre}
\end{table}

\subsection{Problems Encountered and Solutions}

\subsubsection{LVM Volume Group Naming Conflict}

When pre-staging Server-Web's evidence files from the EVE-NG host, the same
volume group conflict from Lab~2
recurred: both the host and the guest disk used \texttt{ubuntu-vg}. The resolution
procedure was identical --- renaming the guest VG to \texttt{helix-web-vg} before
activation --- and the naming rule was formally added to the project's infrastructure
runbook to prevent recurrence in future labs.

\subsubsection{Read-only nbd Device}

On one occasion, the QCOW2 image mounted via \texttt{qemu-nbd} was presented as
read-only despite the \texttt{-readwrite} flag. Investigation revealed that a
previous mount session had not been cleanly disconnected (\texttt{qemu-nbd
\texttt{-{}-disconnect}}), leaving the kernel's NBD client in an inconsistent state.
The fix was to explicitly disconnect any stale NBD device before re-mounting, and
to add a pre-mount check to the project's image manipulation scripts.

\subsubsection{analyst Group Membership for Log Access}

The student account (\texttt{analyst}) was created with restricted permissions so
that it could not read system logs directly as root. However, \texttt{/var/log/auth.log}
requires membership of the \texttt{adm} group on Ubuntu. The pre-staging script was
updated to add \texttt{analyst} to \texttt{adm} at image preparation time, preserving
the educational constraint (no \texttt{sudo} to \texttt{root}) while allowing access
to the evidence files that are the focus of the lab.

\subsubsection{pfSense Console Access via QEMU Monitor}

The pfSense serial console was inaccessible through the standard EVE-NG Telnet
interface due to a configuration mismatch between the QEMU serial port definition
and pfSense's expected console device (\texttt{ttyu0} vs \texttt{ttyS0}). Console
access was recovered by connecting directly to the QEMU monitor socket and issuing
a \texttt{sendkey} sequence to navigate the pfSense boot menu, where the console
port assignment could be corrected persistently.
