% Document: TFG Vol I - Chapter 1 Introduction
% Language: English

\chapter{Introduction}

\section{Problem Statement and Educational Gap}

Cybersecurity education faces a persistent challenge: the gap between theoretical
knowledge of IT systems and the practical understanding of how security failures
manifest in real, integrated environments. Students in computer engineering programmes
learn fundamental concepts in networks, operating systems, databases, and software
architecture --- yet they rarely experience how misconfigurations and design flaws
in these interconnected components translate directly into exploitable vulnerabilities.
Without exposure to multi-component environments that simulate real-world systems,
graduates enter the workforce lacking the diagnostic reasoning and hands-on skills
that professional security roles demand.

This gap has become more acute with the widespread adoption of AI-assisted development
tools. Code generation assistants such as GitHub Copilot and ChatGPT allow developers
to produce functional code rapidly --- authentication logic, input validation routines,
database queries --- without necessarily understanding the security implications of what
is being generated. Vulnerabilities that a security-aware developer would immediately
recognise --- SQL injection sinks, broken access control, unsafe deserialization ---
are routinely introduced through AI-generated code integrated without critical review.
The result is a growing population of practitioners who can build software but cannot
evaluate its attack surface. Practical security education that requires students to
actively exploit and reason about these vulnerabilities --- rather than generate or
copy solutions --- is therefore more urgent than ever.

Within the TecnoCampus Bachelor's Degree in Computer Engineering
(Management and Information Systems track), the \textit{Introduction to Cybersecurity}
course currently lacks a structured, reusable, and automatable lab environment that:

\begin{itemize}
    \item Aligns pentesting scenarios with specific degree learning outcomes
    \item Exposes students to vulnerabilities within integrated, multi-component systems
          rather than isolated challenge-based scenarios
    \item Provides reproducible environments for consistent assessment across cohorts
    \item Reduces instructor workload through automation and structured documentation
    \item Remains cost-effective and institution-independent, deployable on open-source
          infrastructure without vendor lock-in
\end{itemize}

\section{Proposed Solution}

This project proposes a reusable penetration testing lab package built on
EVE-NG~\cite{eve_ng} and deployed on self-hosted infrastructure, designed
specifically for the \textit{Introduction to Cybersecurity} course at TecnoCampus.
The package comprises four progressive lab scenarios: network reconnaissance and
enumeration, web application vulnerabilities, incident response and log forensics,
and privilege escalation. 
%Labs~1--3 are fully implemented and validated; Lab~4 is in progress and will be completed before the submission deadline.
% Commented cuz its inecessary to specify the implementation status in the introduction, and it may change during development.

Each lab is self-contained: it includes a network topology file (\texttt{.unl}),
a student lab guide, an instructor resolution guide with assessment rubric, and
can have supporting automation scripts. Progress is structured around a flag-based model ---
students obtain verifiable tokens at each milestone only by completing the required
technical steps, which prevents superficial shortcuts and ensures that understanding
the attack process is the only path to completion. The entire package is open-source,
vendor-independent, and deployable on any institution running a compatible hypervisor.
Full instructor references and deployment details are available in Volume~II (Appendices)
and at the \href{https://theegea.github.io/TFG/}{project documentation site}.
\begin{comment}
\section{Document Structure}

This report is organised as follows:

\begin{description}
\item[Chapter~1 --- Introduction] Presents the educational gap motivating this
project, the role of AI tools in amplifying security risks, the proposed solution,
and an overview of the document structure.

\item[Chapter~2 --- State of the Art] Reviews the institutional context at
TecnoCampus, existing cybersecurity training platforms, security frameworks,
and prior academic research, identifying the gaps this project addresses.

\item[Chapter~3 --- Objectives and Scope] Defines the project objectives (OE1--OE4),
key performance indicators, and scope boundaries.

\item[Chapter~4 --- Methodology] Describes the iterative development process,
evaluation criteria, and the flag-based progression model adopted for the labs.

\item[Chapter~5 --- Requirements] Specifies functional and non-functional requirements
derived from the project objectives and institutional constraints.

\item[Chapter~6 --- Feasibility Study] Analyses technical, economic, and temporal
feasibility, confirming the viability of the proposed approach within the project timeline.

\item[Chapter~7 --- Development and Implementation] Documents the design decisions,
infrastructure setup, and progressive implementation of Labs~1--4 on the EVE-NG platform.

\item[Chapter~8 --- Conclusions] Evaluates outcomes against the stated objectives,
reflects on lessons learned, and outlines directions for future work.
\end{description}

Supporting materials --- including EVE-NG setup guides, detailed lab instructor references,
and supplementary documentation --- are available in Volume~II (Appendices) and online at
the \href{https://theegea.github.io/TFG/}{project documentation site}.
\end{comment}