% ============================================================
% Autor: Eloi Egea Rada
% Tema: Existing solutions, platforms, and gap analysis
% ============================================================

\chapter{State of the Art}

\section{Context}

The Centre Universitari TecnoCampus~\cite{tecnocampus} offers applied science degrees
oriented towards professional practice in technology, business, creative industries,
and health. Affiliated with the Universitat Pompeu Fabra, it carries that university's
quality seal. Established in 2023 through the merger of three schools --- the Escola
Superior Politecnica (ESUPT), the Escola Superior de Ciencies Socials i de l'Empresa
(ESCSET), and the Escola Superior de Ciencies de la Salut (ESCST) --- it is located
at the technology park of the same name in Mataro. This setting combines academic
infrastructure with an entrepreneurial ecosystem that supports student involvement in
innovation and industry-facing projects.

The primary beneficiary of this project is the Bachelor's Degree in Computer
Engineering (Management and Information Systems track). Students in this programme
build foundational skills in networks, operating systems, databases, and software
engineering --- all of which are directly relevant to security practice.
The \textit{Introduction to Cybersecurity} course is a natural integration point
for these skills, yet it currently lacks a dedicated hands-on environment.

Three perspectives illustrate the gap this project addresses:

\paragraph{Institutional perspective.}
The absence of dedicated cybersecurity lab infrastructure limits the institution's
capacity to offer competitive, hands-on security training aligned with degree learning
outcomes. Commercial alternatives such as HackTheBox and TryHackMe are externally
hosted, non-customisable, and not designed to map to specific academic curricula or
assessment rubrics.

\paragraph{Student perspective.}
Students encounter security concepts as isolated theoretical constructs rather than
as failures observable in realistic, integrated environments. Without exposure to
multi-component systems --- firewalls, routers, servers, and applications working
together --- they cannot develop the diagnostic reasoning required in professional
security roles.

\paragraph{Teacher perspective.}
Designing and maintaining practical security exercises is resource-intensive.
The absence of reusable, automatable lab environments, teachers must improvise
or rely on third-party platforms, which reduces pedagogical control and limits
reproducibility across cohorts.



\section{Existing Platforms Overview}

The major platforms used for cybersecurity training share a common limitation: they are designed
for independent learners or certification preparation rather than curriculum-aligned institutional
deployment. Table~\ref{tab:platform_comparison} summarises their key characteristics.
Platforms such as HackTheBox~\cite{hackthebox} and TryHackMe~\cite{tryhackme} are externally hosted
and cannot be deployed on institutional infrastructure. SEED Labs~\cite{seedlabs} and
GOAD~\cite{goad} support local deployment but cover only isolated concepts or narrow domains.
VulnHub~\cite{vulnhub} and Hack4u~\cite{hack4u} lack automation and assessment integration.
Cisco Networking Academy~\cite{cisco_netacad} provides formal structure but is tightly coupled
to vendor technology and limited to network simulation.

\begin{table}[H]
\centering
\scriptsize
\setlength{\tabcolsep}{4pt}
\begin{tabularx}{\textwidth}{lXXXXX}
\toprule
\textbf{Platform} & \textbf{Approach} & \textbf{Deploy} & \textbf{Curriculum} & \textbf{Auto} & \textbf{Network} \\
\midrule
HackTheBox & CTF & No & Minimal & No & Isolated \\
TryHackMe & Gamified & No & Structured & No & Limited \\
Cisco Academy & Academic & Yes (PT) & Strong & Limited & Net-focused \\
SEED Labs & Conceptual & Yes (Docker) & Concept-based & Minimal & Isolated \\
GOAD & AD-focused & Yes (\mbox{Terraform}) & None & Moderate & \mbox{AD Domain} \\
VulnHub & Community & Yes & None & No & Isolated \\
Hack4u Academy & Methodology & External & Strong & Minimal & Limited \\
\bottomrule
\end{tabularx}
\caption{Comparative Analysis: Existing Cybersecurity Training Platforms}
\label{tab:platform_comparison}
\end{table}

Full individual platform descriptions, including detailed discussion of each platform's
pedagogical approach, deployment model, and specific limitations, are provided in
the Appendix~E.

\section{Security Frameworks and Methodologies}

Several authoritative frameworks inform cybersecurity education and guide both attack
methodologies and defensive best practices.

\subsection{PTES \textendash{} Penetration Testing Execution Standard}

The Penetration Testing Execution Standard (PTES)~\cite{ptes} defines a complete methodology
for professional penetration testing, structured in seven phases: pre-engagement interactions,
intelligence gathering, threat modelling, vulnerability analysis, exploitation, post-exploitation,
and reporting. PTES provides the overarching procedural framework for this project: Lab~1
(reconnaissance and enumeration) maps to the intelligence gathering and vulnerability analysis
phases, while Labs~2--4 progressively cover exploitation and post-exploitation techniques.

\subsection{OWASP Top 10}

The OWASP Top 10 is a regularly updated list of the most critical web application security
vulnerabilities \cite{owasp2021}. For educational purposes, it provides a focused, industry-recognised
curriculum that prioritises the attack vectors students are most likely to encounter in practice.
Lab 2 of this project is directly aligned with OWASP Top 10 categories.

\subsection{CIS Critical Controls}

The CIS Critical Controls (v8.0) represent a prioritised set of defensive techniques spanning
network defence, data protection, and incident response \cite{cis_controls}. Labs 1 and 3
incorporate defensive controls so that students understand not only how to attack but why
specific security measures are implemented.

\subsection{NIST Cybersecurity Framework}

The NIST Cybersecurity Framework organises security activities into five functions: Identify,
Protect, Detect, Respond, and Recover \cite{nist_framework}. The lab design follows these
principles by enabling students to identify vulnerabilities through hands-on exercises and
understand corresponding protective measures.




\section{Prior Academic Research}

Academic research consistently validates the effectiveness of hands-on, laboratory-based
approaches in cybersecurity education. Conti and Di Pietro \cite{conti2019} show that
challenge-based environments significantly improve skill retention and practical competency
compared to purely theoretical instruction. Murphy et al.\ \cite{murphy2018} provide a
structured methodology for designing laboratory exercises explicitly aligned with curriculum
outcomes, validating the need for scenario-specific tasks and integrated assessment.
Sponj and Heri\v{c}ko \cite{sponj2021} conducted a systematic review of cybersecurity
education research and identified hands-on practice and authentic task design as the two
most influential factors in learner engagement and competency development.
Al~Mohanadi and Furnell \cite{almohannadi2016} establish a requirements-driven approach to
information security assessment laboratories, emphasising that lab tasks must map directly
to defined learning outcomes to achieve meaningful pedagogical impact.

These findings directly inform the design choices in this project: each lab is built around
scenario-based flags mapped to specific learning objectives, includes an teacher reference
sheet, and culminates in an automated assessment mechanism. Together, these elements address
the gap between platform availability and curriculum integration identified in the literature.

\section{Gap Analysis: What Existing Solutions Miss}

Despite the proliferation of cybersecurity training platforms, significant gaps remain:

\textbf{Gap 1: Curriculum Integration.}
Most platforms target independent learners or certification preparation rather than integration
into specific degree programmes. No major platform explicitly maps pentesting scenarios to the
learning outcomes of a particular computer science curriculum.

\textbf{Gap 2: Institution Control.}
Effective educational deployment requires institutional control over content, assessment, and
infrastructure. External platforms introduce dependencies and limit customisation, while
fully self-hosted solutions often impose high maintenance burdens.

\textbf{Gap 3: Integrated Network Penetration Testing.}
Most platforms focus on isolated services or individual techniques. Few provide realistic,
multi-system network topologies where reconnaissance informs a complete attack chain and
lateral movement is required.

\textbf{Gap 4: Automation and Scalability.}
Even classroom-oriented platforms often lack automated deployment, environment reset, and
assessment mechanisms, which increases teacher workload and limits large-scale institutional
adoption.

\textbf{Gap 5: Balance of Guidance and Autonomy.}
Effective labs must balance sufficient structure to keep students on track with enough
openness to encourage problem-solving and methodology development.


\section{The Need for EVE-NG Based Labs}

This analysis shows that while excellent individual components exist---gamified learning
(TryHackMe), challenging scenarios (HackTheBox), research-backed content (SEED Labs),
specialised domains (GOAD)---no existing solution comprehensively addresses the needs of a
university cybersecurity education programme:

\begin{itemize}
    \item \textbf{Curriculum-aligned labs}: Mapped to specific degree learning outcomes
    \item \textbf{Integrated network scenarios}: Multi-system penetration testing with realistic complexity
    \item \textbf{Institutional control}: Locally deployable, customisable, and maintainable
    \item \textbf{Complete automation}: One-command deployment and reset with built-in assessment
    \item \textbf{System-level realism}: Full virtual machine workflows (real OS images, ISO-based
          installation, native kernel behaviour) that reproduce the technical conditions of
          professional pentesting practice
\end{itemize}

This bachelor's thesis bridges that gap by using EVE-NG \cite{eve_ng} to create a
comprehensive, reusable, and automatable teaching package for cybersecurity practical labs.
It draws on the pedagogical strengths of existing platforms while addressing their limitations
through institutional control, curriculum alignment, and automation.


% END OF CHAPTER 2: STATE OF THE ART
