% ============================================================
% Autor: Eloi Egea Rada
% Tema: Research methodology, development phases, and validation approach
% ============================================================

\chapter{Methodology}
\label{ch:methodology}

This chapter outlines the research strategy, development approach, validation methods, and
project lifecycle management employed in this bachelor's thesis. The project integrates
principles from software engineering, cybersecurity pedagogy, and empirical research to
ensure systematic development and validation of the EVE-NG\cite{eve_ng}-based teaching package.

\section{Research and Development Phases}

The project is structured in four distinct phases:

\subsection{Phase 0: Preparation and Infrastructure (July--November 2025)}

\textbf{Objective}: Establish the technical foundation and personal capacity for the project.
Activities included HomeLab infrastructure initialisation (Proxmox\cite{proxmox}), development
environment setup (VSCode\cite{vscode}, LaTeX\cite{latex_project}, Git), and security tools
installation (Parrot OS\cite{parrot_os}, Metasploit\cite{metasploit}, Burp Suite\cite{burp_suite}).

\textbf{Duration}: 105 hours \quad \textbf{Status}: Completed (100\%)

\subsection{Phase 1: Preliminary Project Proposal and Conceptual Design (October 2025--January 2026)}

\textbf{Objective}: Define project scope, objectives, and detailed planning.
Work covered SMART objectives definition, state-of-the-art analysis of 7 platforms,
functional and non-functional requirements specification, pentesting methodology selection
(PTES\cite{ptes}, OWASP\cite{owasp_testing_guide}, CEH\cite{ceh_framework}), risk analysis
with 12+ identified risks, resource allocation, and Gantt chart construction with 37 tasks
across the critical path.

\textbf{Duration}: 188 hours \quad \textbf{Status}: Completed \quad
\textbf{Milestone}: January 16, 2026

\subsection{Phase 2: Implementation and Interim Validation (January--March 2026)}

\textbf{Objective}: Develop the first three laboratories and conduct preliminary validation.
Labs 1--3 were built and tested (reconnaissance, web vulnerabilities, and network/incident
response), covering research, topology design, vulnerability implementation, and teaching
material production. Informal pilot sessions with 4--5 GEISI students were conducted as each
lab was completed, with usability observations, post-session surveys, and feedback-driven
adjustments.

\textbf{Duration}: 212 hours \quad \textbf{Status}: In progress \quad
\textbf{Milestone}: March 20, 2026

\subsection{Phase 3: Completion and Final Validation (March--May 2026)}

\textbf{Objective}: Develop Lab 4, complete automation scripts, and conduct final validation.
Activities include privilege escalation lab development (CEH framework), two rounds of
comprehensive penetration testing across all four labs, utility script development (Bash/Python),
final pilot validation with the student group, and full bachelor's thesis report writing and
revision.

\textbf{Duration}: 315 hours \quad \textbf{Status}: In progress \quad
\textbf{Final Deadline}: May 27, 2026

\section{Information Search Strategies}

Research draws on academic databases (IEEE Xplore, ACM Digital Library, Scopus, Google Scholar)
and primary industry standards: PTES\cite{ptes}, OWASP\cite{owasp_testing_guide},
NIST\cite{nist_framework,nist_sp800115}, CIS Controls\cite{cis_controls}, and
EC-Council CEH\cite{ceh_framework}. Platform and tool documentation was consulted for
EVE-NG\cite{eve_ng_docs}, Parrot OS\cite{parrot_os}, Metasploit\cite{metasploit},
OWASP ZAP\cite{owasp_zap}, and Burp Suite\cite{burp_suite}. Primary research uses informal
pilot sessions with GEISI students, post-session surveys, and quantitative metrics
(time-on-task, completion rates).

\section{Development Methodology}

The project adopts an iterative, agile-inspired approach:

\begin{itemize}
  \item \textbf{Design}: Conceptual architecture and topology planning
  \item \textbf{Implementation}: Lab creation, vulnerability implementation, tooling
  \item \textbf{Validation}: Security testing, educational effectiveness review
  \item \textbf{Iteration}: Feedback incorporation and refinement
  \item \textbf{Documentation}: Technical and pedagogical material preparation
\end{itemize}

All artefacts are version-controlled in a GitHub repository. Formal documentation is produced
in LaTeX (this memory, Volume~I) and published as a MkDocs web site for operational reference.
Instructor deployment references, startup checklists, and credential tables are compiled
separately in the Annexos (Volume~II).

\section{Platform Selection Rationale}

A detailed comparative analysis of existing cybersecurity training platforms ---
including HackTheBox, TryHackMe, Cisco NetAcad, SEED Labs, GOAD, VulnHub, and Hack4u ---
is presented in the State of the Art chapter. That analysis evaluates each platform across
customisation, institutional control, automation, cost, and curriculum alignment, and
concludes that EVE-NG is the optimal choice. Its key advantages --- open-source licensing,
full topology customisation, local deployability, and automation support --- directly address
the gaps identified in existing platforms.

\section{Validation and Success Criteria}

Success is measured against the KPIs defined in Chapter~4, covering three dimensions:
technical performance (deployment time, boot success rate, reproducibility), educational
effectiveness (lab resolution time, completion rate, user satisfaction), and sustainability
(documentation coverage, open-source compliance, institutional readiness). Validation is
conducted through pilot sessions with a representative group of GEISI students and through
automated technical checks at each development milestone.

\section{Risk Management}

\begin{table}[H]
\centering
\small
\begin{tabularx}{\textwidth}{lXl}
\toprule
\textbf{Risk} & \textbf{Mitigation Strategy} & \textbf{Owner} \\
\midrule
Scope creep / Feature additions & Clear scope boundaries and change management & Eloi \\
EVE-NG compatibility issues & Extensive testing on Community and Pro editions & Eloi \\
Pilot user dropout & Clear incentives and scheduling flexibility & Pere + Eloi \\
Performance degradation & Early benchmarking and optimisation iterations & Eloi \\
Documentation gaps & Peer review and user feedback cycles & Eloi + Faculty \\
Institutional adoption barriers & Change management plan and training support & Pere \\
\bottomrule
\end{tabularx}
\caption{Key project risks and mitigation strategies}
\label{tab:risk_management}
\end{table}

Full phase activity lists, quality assurance procedures, and documentation
practices are provided in Appendix~\ref{app:methodology-detail}.
