% ============================================================
% Autor: Eloi Egea Rada
% Tema: Conclusions, outcomes evaluation, and future work
% ============================================================

\chapter{Conclusions}
\label{ch:conclusions}

This chapter evaluates the outcomes of the project against the stated objectives,
reflects on the lessons learned during development, and outlines directions for
future work. The overall goal of the project --- to design and implement a reusable,
curriculum-aligned penetration testing lab package for the \textit{Introduction to
Cybersecurity} course at TecnoCampus --- has been substantially achieved.

\section{Summary of Achievements}

The project delivered four progressive lab scenarios built on EVE-NG~\cite{eve_ng}
and deployed on a self-hosted Proxmox~\cite{proxmox} infrastructure:

\begin{itemize}
    \item \textbf{Lab~1 --- Reconnaissance and Enumeration}: A layered network
          topology featuring pfSense~\cite{pfsense}, VyOS~\cite{vyos}, and an
          Ubuntu Server~\cite{ubuntu} target. Students perform passive and active
          reconnaissance using Nmap, DNS enumeration, and service fingerprinting,
          mirroring the PTES~\cite{ptes} initial access phase.

    \item \textbf{Lab~2 --- Web Application Vulnerabilities (SYNAPSE)}: A custom
          two-tier web application aligned with OWASP Top~10~\cite{owasp2021}.
          The exploitation chain --- stored XSS, SQL injection, and command injection
          --- guides students through a realistic attack scenario requiring lateral
          thinking and multi-step reasoning.

    \item \textbf{Lab~3 --- Incident Response and Log Forensics (HELIX Systems)}:
          An inverted-role lab where students act as security analysts following NIST
          SP~800-61~\cite{nist_sp80061}. Pre-staged log evidence and a reconstructed
          attack timeline develop forensic analysis skills alongside offensive concepts.

    \item \textbf{Lab~4 --- Privilege Escalation}: Under development (Phase~3). The
          topology and scenario design are complete; implementation and validation
          will be finalised before the May~27, 2026 submission deadline.
\end{itemize}

Supporting each lab is a student guide, an instructor resolution guide with assessment
rubric, and automation scripts for deployment and environment reset. All documentation
is published in three parallel layers: the formal LaTeX report (this volume),
Volume~II (Annexos), and the project web documentation site at
\href{https://TheEgea.github.io/TFG/}{TheEgea.github.io/TFG}.

\section{Objectives Evaluation}

Table~\ref{tab:objectives_evaluation} summarises the four specific objectives
defined in Chapter~\ref{ch:objectives} and their current completion status.

\begin{table}[H]
  \centering
  \small
  \setlength{\tabcolsep}{4pt}
  \renewcommand{\arraystretch}{1.2}
  \begin{tabularx}{\textwidth}{l X l}
    \toprule
    \textbf{Objective} & \textbf{Description} & \textbf{Status} \\
    \midrule
    OE1 & Design and implement 4 EVE-NG labs     & 75\% (3/4 completed) \\
    OE2 & Develop automation scripts              & Completed \\
    OE3 & Develop structured teaching materials   & 75\% (3/4 labs documented) \\
    OE4 & Validate usability and effectiveness    & Partial (pilot pending) \\
    \bottomrule
  \end{tabularx}
  \caption{Specific objectives evaluation at Phase~3 mid-point}
  \label{tab:objectives_evaluation}
\end{table}

\textbf{OE1} is on track: Labs~1--3 are fully implemented and validated.
Lab~4 design is complete; implementation is the primary remaining activity in
Phase~3. \textbf{OE2} is fully achieved: automation scripts cover ISO upload,
bridge configuration, and connectivity checks, with at least one utility script
per lab as required. \textbf{OE3} follows OE1 progress --- three complete student
and instructor guide pairs are available; Lab~4 materials will be produced in
parallel with implementation. \textbf{OE4} remains partially open: informal
pilot observations conducted after Labs~1--3 confirmed usability and flag
reachability, but formal KPI measurement (time-on-task, satisfaction survey)
is reserved for the final pilot session in May~2026.

\section{Lessons Learned}

\paragraph{Infrastructure stability precedes content.}
The migration from VMware to Proxmox early in Phase~0 was the single most impactful
decision of the project. Attempting to develop lab content on an unstable hypervisor
would have introduced compounding delays. The lesson is general: validate the
execution environment thoroughly before investing in content.

\paragraph{Image management is a project in itself.}
Selecting, importing, and maintaining QEMU images for EVE-NG required significantly
more effort than anticipated. Naming conventions, permission scripts, and QCOW2
format validation each required dedicated attention. The \texttt{Iso\_uploader.py}
script emerged from this experience and reduced future onboarding time substantially.

\paragraph{Flag-based progression is effective for self-paced learning.}
The flag model --- where students collect verifiable tokens at each milestone ---
provides immediate feedback and natural progress checkpoints without requiring
instructor presence. Informal pilot observations confirmed that students find the
model motivating and that the flags help them self-diagnose when they are stuck
on the wrong path.

\paragraph{Documentation as a first-class deliverable.}
Producing documentation in three layers (LaTeX, web, src/) introduced overhead
but proved valuable: the web documentation became the primary reference during
informal pilot sessions, while the LaTeX volume serves the formal academic record.
Future projects of this type should plan the documentation architecture before
beginning implementation.

\paragraph{Time estimates for realistic EVE-NG scenarios are consistently optimistic.}
Each lab required approximately 20--30\% more hours than the initial estimate,
primarily due to unexpected node configuration persistence issues and the need to
validate flag reachability after each topology change. Build in explicit buffer
for integration testing in any similar project plan.

\section{Limitations and Future Work}

\subsection{Current Limitations}

\begin{itemize}
    \item \textbf{Lab~4 not yet complete}: Privilege escalation content is under
          development. The full four-lab package will be available upon submission.

    \item \textbf{Formal pilot not yet conducted}: KPI validation (time-on-task,
          satisfaction score, completion rate) requires the scheduled May~2026 pilot
          session with GEISI students. Results will be incorporated into the final
          report if completed before the submission deadline.

    \item \textbf{No LMS integration}: The package is self-contained but not
          integrated with Moodle or any institutional learning management system.
          Assessment rubrics are PDF-based; submission and grading are manual.

    \item \textbf{Single-group concurrency}: The current infrastructure supports one
          lab group at a time. Running simultaneous sessions for multiple groups would
          require either additional server resources or a cloud-based EVE-NG deployment.
\end{itemize}

\subsection{Directions for Future Work}

\begin{itemize}
    \item \textbf{Complete Lab~4}: Finalise privilege escalation scenario with SUID,
          sudo misconfiguration, and cron-based persistence vectors; produce student
          guide and instructor reference.

    \item \textbf{Expand the OWASP coverage}: Lab~2 currently covers five Top~10
          categories. Additional modules could address server-side request forgery
          (SSRF), insecure deserialization, and API security misconfigurations.

    \item \textbf{Moodle integration}: Export assessment rubrics as Moodle-compatible
          grading forms; implement automated flag submission via a simple REST endpoint.

    \item \textbf{Multi-group support}: Evaluate cloud EVE-NG deployment (AWS or
          on-premises scale-out) for concurrent execution by multiple student groups
          within the same course edition.

    \item \textbf{Lab~5 --- Active Directory}: A Windows domain penetration testing
          scenario (Kerberoasting, Pass-the-Hash, lateral movement) to complement
          the existing Linux-centric labs and address OSCP-aligned skill gaps.
\end{itemize}

\section{Personal Reflection}

This project brought together skills from across the Computer Engineering degree ---
network administration, operating system internals, web security, scripting, and
technical writing --- within a single coherent artefact. The process of designing
realistic attack scenarios, implementing them in a reproducible environment, and
documenting them rigorously to serve as teaching material was more challenging and
more rewarding than anticipated.

Working within the constraints of academic infrastructure and a fixed timeline
required constant prioritisation. The decisions that proved most durable were those
made at the architectural level early in Phase~0: the choice of Proxmox over VMware,
EVE-NG Community Edition over proprietary alternatives, and an open-source-only tool
stack. These decisions ensured that the package remains deployable and maintainable
by any institution with comparable hardware, independent of licensing or vendor
relationships.

The experience reinforced a conviction that practical, scenario-based learning is
essential for cybersecurity education, and that the gap between theoretical knowledge
and professional practice can be narrowed significantly through well-designed lab
environments. Delivering a package that will be used in a real course is the most
meaningful validation of the work.
