\chapter*{Dedication and Acknowledgements}
\addcontentsline{toc}{chapter}{Dedication and Acknowledgements}

\section*{Dedication}

This work is dedicated to all students and professionals interested in cybersecurity,
particularly in ethical pentesting.
We hope that the labs and scenarios presented in this bachelor's thesis will be useful
for their training and professional development.

\section*{Acknowledgements}

\subsection*{Institutional Acknowledgements}
I would like to express my sincere gratitude to my tutor, Pere Vidiella i Catalan,
for his guidance, trust, and constructive feedback throughout this project.
I am also deeply grateful to the faculty and staff of the Computer Engineering program
for providing the academic framework, resources, and supportive environment that made
this work possible. Their dedication to excellence in teaching and mentoring has been
instrumental in shaping both the technical scope and the academic rigor of this thesis.

\subsection*{Personal Acknowledgements}
This thesis is dedicated, above all, to my grandmother,
who passed away just before I began my degree---shortly after my university admission
and the day before my English examination. Choosing to write this thesis in English is,
in part, a meaningful way to close a chapter that marked the beginning of my university
journey without the person who meant the most to me and who, in many ways, was like
a mother to me.

I am also profoundly grateful to my friends, my partner, and my partner's family for
their unwavering support and encouragement, especially during the most difficult moments
of my studies, including a period when I faced a serious health challenge that could have
left me paralyzed. Their presence, care, and belief in me have been essential to completing
this work.

\subsection*{Use of Artificial Intelligence Tools}
During the development of this thesis, complementary AI tools were used exclusively for
review and error-detection purposes: a locally-hosted language model (LM~Studio) and
an online assistant (Claude, Anthropic). These tools were not used to generate any
academic content; all analysis, design, implementation, and writing are the author's own work.

\section*{Personal Reflection}
The completion of this Bachelor's Thesis has been an enriching experience that has allowed
me to deepen my knowledge of cybersecurity and develop technical and project management skills.
I have learned the importance of practice in learning this discipline and how virtualized
environments can facilitate this process.
This project has motivated me to continue exploring the field of cybersecurity and to
contribute to the training of future professionals in this critical area.
\newpage
